---
title: "STATS 250 Lab 6 Report"
author: "YOUR NAME GOES HERE IN QUOTES"
date: "TODAY'S DATE GOES HERE IN QUOTES"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    theme: lumen
    css: assets/styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("shuffle_two_groups.R") # This imports the shuffle_two_groups() function
set.seed(42) # Note setting the seed!!
```

# Instructions

Welcome to Lab 6! Complete the following exercises designed to help you understand the simulation process for making inference about a population proportion. It is important that you keep your answers inside the ".answer" blocks. Do NOT delete the three colon blocks (:::) and keep your responses between them. Use the tutorial for help, and feel free to ask questions on Piazza or email your lab instructor.

Depending on your instruction method (synchronous or asynchronous), complete the following exercises collaboratively.

1. **Synchronous remote:** Work with your peers in breakout rooms during the synchronous lab.
1. **Asynchronous remote:** Use Piazza for asynchronous collaboration.

**Collaborators:** If applicable, replace this text with the names of anyone you collaborated with on this project.

**NOTE:** If you perform a calculation in R and assign it a name but do not print the value in your knitted document, we reserve the right to deduct points.

# Background

:::{.background}
Are students who transfer to U-M more likely to be from Michigan than non-transfer students?
Imagine you're helping the Registrar's office to answer this question using data from the STATS 250 Student Survey.

We took a random sample of 200 STATS 250 students who responded to the Student Survey, which asked them to tell us whether they were a transfer student, as well as their residency status (in-state vs. out-of-state/international).
**Of the 200 sampled students, 32 transferred to U-M. 22 of those transfer students are in-state. 119 students in the sample are in-state.**

Use this information to answer the below questions to address our research question: **"Is the proportion of transfer students who are from in state greater than the proportion of non-transfer students who are from in state, based on the information from the Stats 250 student survey?"**
:::

# Lab Assignment

**Don't forget to run the setup chunk on line 14.**

## Question 1
:::{.question}
State the hypotheses we want to test *using symbols*.
:::

:::{.answer}
- $H_0$: *Replace this text with your null hypothesis*
- $H_A$: *Replace this text with your alternative hypothesis*
:::

## Question 2
:::{.question}
Start by identifying our groups and define a success and a failure.
:::

:::{.answer}
- **Group A** is
- **Group B** is
- A **success** is
- A **failure** is 
:::

## Question 3
::: {.question}
Start by tabulating our observed data. Fill in the table below based on the information given in the background above.
:::

::: {.answer}

|             | Group A | Group B | Total |
|:-----------:|:-------:|:-------:|:-----:|
| **Failure** |         |         |       |
| **Success** |         |         |       |
| **Total**   |         |         |       |

:::

## Question 4
:::{.question}
Build the data frame that is represented by your table above. Call this data frame `stats250_students`.
:::

:::{.answer}
```{r question4, error = T}
# Replace this comment with code required for Question 4. (Remember that this text is a comment, so R ignores it; you can delete it if you want.)

```
:::

## Question 5

:::{.question}
Make a table of your newly created data frame `stats250_students` to verify that the table matches your answers in Question 3. If your table does not match, go back and rework Question 4 until it does match.
:::

:::{.answer}
```{r question5, error = T}
# Replace this comment with code required for Question 5. (Remember that this text is a comment, so R ignores it; you can delete it if you want.)

```
:::

## Question 6
:::{.question}
Compute the observed value of the statistic, $\hat{p}_1 - \hat{p}_2$ using R as a calculator. 

Remember that the syntax in between dollar signs in R Markdown is a way to make math look nice, not R syntax. If you give the value a name, don't forget to print the object so the value appears in your knitted document.
:::

:::{.answer}
```{r question6, error = T}
# Replace this comment with code required for Question 6. (Remember that this text is a comment, so R ignores it; you can delete it if you want.)

```

:::

## Question 7
:::{.question}
Use the `replicate()` function to run the simulation 100 times, making sure to assign your results a name. 
:::

:::{.answer}
```{r question7, error = T}
# Replace this comment with code required for Question 7. (Remember that this text is a comment, so R ignores it; you can delete it if you want.)

```
:::

## Question 8 
:::{.question}
Make a histogram of your simulation results from Question 7, with an appropriate title and axis labels. Draw a vertical line at the observed value of the statistic, $\hat{p}_1 - \hat{p}_2$ that was found in Question 6.
:::

:::{.answer}
```{r question8, error = T}
# Replace this comment with code required for Question 8. (Remember that this text is a comment, so R ignores it; you can delete it if you want.)

```
:::

## Question 9
:::{.question}
Compute the p-value, the proportion of simulated results as extreme or more extreme (in the direction of the alternative hypothesis) than the observed value of the statistic, $\hat{p}_1 - \hat{p}_2$.
:::

:::{.answer}
```{r question9, error = T}
# Replace this comment with code required for Question 9. (Remember that this text is a comment, so R ignores it; you can delete it if you want.)

```
:::

# Wrap-Up and Submission

At the top of the document, make sure you've changed the `author` field to your name (in quotes!). If you'd like, change the date as well.

When you've finished the lab, click the **Knit** button one last time.

## Submission instructions

### If you're using RStudio Cloud

1.  In the Files pane, check the box next to `lab06-report.html`.
2.  Click More \> Export...
3.  Click Download and save the file on your computer in a folder you'll remember and be able to find later. You can just call the file `lab06-report.html` or whatever else you'd like (as long as you remember what you called it).

### If you're using RStudio installed on your computer

1.  locate the `lab06-report.html` file on your computer. The file will be saved in the location indicated at the top of the files pane.

### Submission to Canvas

1.  Click the "Assignments" panel on the left side of the page. Scroll to find "Lab 6", and open the assignment. Click "Submit Assignment".
2.  Towards the bottom of the page, you'll be able to choose `lab06-report.html` from the folder you saved it in from RStudio Cloud or noted if you're using RStudio Desktop. **You will only be able to upload a .html file -- do not upload any other file type.**
3.  Click "Submit Assignment". You're done!
